Towards typed repositories of proofs

### 2010-06-18 papers/MIPS2010/abstract.tex
* Derniere passe sur intro.
git-svn-id: http://gasp.dev.regis-gianas.com/svn/trunk@128 403e655e-980a-47bd-8c60-1ee5348bb6b7

We believe that these tools are not adapted for the new, demanding
requirements of proof developments. Indeed, whereas compilation of a
program is usually fast enough for the programmer to rely on the usual
interaction loop ((edit; compile)*; commit)*, the operation of proof
checking is usually too expensive computationally to mimic this
workflow. But even besides the time factor, this ``traditional'' way
of formalizing mathematics hinders the process of mathematical
discovery process: once a concept contained in a file is compiled, it
is considered frozen and any changes to it require the recompilation
of the whole file; the linearity of the development also gives no room
for alternate, inequivalent definitions. This fact has nonetheless
been shown to be of crucial to the mathematical discovery
process~\citep{lakatos1964proofs}, and we believe that they should be
taken into account in the realization of mathematical assistants.

In fact, although dedicated tools exist to formalize the description
of languages and their metatheory (e.g. \textsf{Twelf},
{\citep{pfenning1999system}}), and substantial formalizations have been
undertaken~\citep{lee07}, we still use legacy tools based on text
representation to manage our developments. The general goal exposed
here is to replace this tool chain and make it language-aware, both on
a syntactic side through the use of abstract syntax trees (AST) instead of
concrete syntax and on a semantic side by using typing to ensure 
repository consistency.

>>>>

We believe that these tools are not adapted for the new, demanding
requirements of proof developments. Indeed, whereas compilation of a
program is usually fast enough for the programmer to rely on the usual
interaction loop ((edit; compile)*; commit)*, the operation of proof
checking is usually too expensive computationally to mimic this
workflow. But even beyond the time factor, this ``traditional'' way
of formalizing mathematics hinders the process of mathematical
discovery process: once a concept contained in a file is compiled, it
is considered frozen and any changes to it require the recompilation
of the whole file; the linearity of the development also gives no room
for alternate, inequivalent definitions. This fact has nonetheless
been shown to be crucial to the mathematical discovery
process~\citep{lakatos1964proofs}, and we believe that they should be
taken into account in the realization of mathematical assistants.

In fact, although dedicated tools exist to formalize the description
of languages and their metatheory (e.g. \textsf{Twelf},
{\citep{pfenning1999system}}), and substantial formalizations have been
undertaken~\citep{lee07}, we still use legacy tools based on text
representation to manage our developments. The general goal exposed
here is to replace this tool chain and make it language-aware, both on
a syntactic side through the use of abstract syntax trees (AST) instead of
concrete syntax, and on a semantic side by using typing to ensure 
repository consistency.
